---
title: "Assignment 2: Coding Basics"
author: "Leyi Ruan"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: console
---

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on coding basics. 

## Directions
1. Change "Student Name" on line 3 (above) with your name.
2. Work through the steps, **creating code and output** that fulfill each instruction.
3. Be sure to **answer the questions** in this assignment document.
4. When you have completed the assignment, **Knit** the text and code into a single PDF file.
5. After Knitting, submit the completed exercise (PDF file) to the dropbox in Sakai. Add your first and last name into the file name (e.g., "FirstLast_A02_CodingBasics.Rmd") prior to submission.



## Basics Day 1

1. Generate a sequence of numbers from one to 100, increasing by fours. Assign this sequence a name. 

2. Compute the mean and median of this sequence. 

3. Ask R to determine whether the mean is greater than the median. 

4. Insert comments in your code to describe what you are doing.

```{r}
#1. I'm building a sequence from 1 to 100, increased by 4. 
#Then assign the sequence just built to seq1.
seq1 <- seq(1, 100, 4)
seq1

#2. Use mean command to compute the mean of the sequence, 
#and use median command to calculate the median of the sequence.
mean1 <- mean(seq1)
mean1 
median1 <-median(seq1)
median1

#3. Use conditional statements to check if mean is greater than median. 
#The result turns out to be false. Thus, mean is not greater than median.
mean1 > median1
```

## Basics Day 2

5. Create a series of vectors, each with four components, consisting of (a) names of students, (b) test scores out of a total 100 points, and (c) whether or not they have passed the test (TRUE or FALSE) with a passing grade of 50.

6. Label each vector with a comment on what type of vector it is.

7. Combine each of the vectors into a data frame. Assign the data frame an informative name.

8. Label the columns of your data frame with informative titles.

```{r}
# 5 & 6. Create 3 vectors: names, scores, and results
names <- c("Lily", "Mark", "James", "Bob") #Character Vector
scores <- c(99, 68, 35, 51) #Integer vector
results <- c(TRUE,TRUE, FALSE, TRUE) #Logical vector

# 7 . Combine vector into a dataframe
df_names <- as.data.frame(names)
df_scores <- as.data.frame(scores)
df_results <- as.data.frame(results)
test <- cbind(df_names, df_scores, df_results)
test

# 8. Label Columns with Informatives Titles
test <- data.frame("Name"=names, "Score"=scores, "Result"=results)
test
```

9. QUESTION: How is this data frame different from a matrix?

> Answer: In a data frame the columns contain different types of data, but in a matrix all the elements are the same type of data.

10. Create a function with an if/else statement. Your function should determine whether a test score is a passing grade of 50 or above (TRUE or FALSE). You will need to choose either the `if` and `else` statements or the `ifelse` statement. Hint: Use `print`, not `return`. The name of your function should be informative.

11. Apply your function to the vector with test scores that you created in number 5. 

```{r}
# 10. Create Function to examine Pass/Fail
Pass_Fail <- function(x){
  if(x>=50) {
    Result_2 <- "TRUE"
  }
  else{
    Result_2 <- "FALSE"
  }
  return(Result_2)
}

# 11. Apply the function to the vector with test scores
# To distinguish the result derived from using function with the original #result vector, I name the new list derived from lapply as Result_2
Result_2 <- lapply(scores, Pass_Fail)
Result_2 #Result_2 is a list
```

12. QUESTION: Which option of `if` and `else` vs. `ifelse` worked? Why?

> Answer: I think theoretically both should work, but I'm not sure how to return values from ifelse statement. Thus, if else statement works better under this scenario.
