---
title: "Assignment 5: Data Visualization"
author: "Leyi Ruan"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: console
---

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on Data Visualization 

## Directions
1. Change "Student Name" on line 3 (above) with your name.
2. Work through the steps, **creating code and output** that fulfill each instruction.
3. Be sure to **answer the questions** in this assignment document.
4. When you have completed the assignment, **Knit** the text and code into a single PDF file.
5. After Knitting, submit the completed exercise (PDF file) to the dropbox in Sakai. Add your last name into the file name (e.g., "Fay_A05_DataVisualization.Rmd") prior to submission.

The completed exercise is due on Tuesday, February 23 at 11:59 pm.



## Set up your session 
1. Set up your session. Verify your working directory and load the tidyverse 
and cowplot packages. Upload the NTL-LTER processed data files for nutrients 
and chemistry/physics for Peter and Paul Lakes (both the tidy 
[`NTL-LTER_Lake_Chemistry_Nutrients_PeterPaul_Processed.csv`] and the gathered
[`NTL-LTER_Lake_Nutrients_PeterPaulGathered_Processed.csv`] versions) and 
the processed data file for the Niwot Ridge litter dataset. 
```{r,warning = FALSE}
getwd()
library(tidyverse)
library(cowplot)
PeterPaul.chem.nutrients <- 
  read.csv("../Data/Processed/NTL-LTER_Lake_Chemistry_Nutrients_PeterPaul_Processed.csv", stringsAsFactors = TRUE)
PeterPaul.chem.nutrients.gathered <-
  read.csv("../Data/Processed/NTL-LTER_Lake_Nutrients_PeterPaulGathered_Processed.csv", stringsAsFactors = TRUE)
Niwot_Ridge.sum <-
  read.csv("../Data/Processed/NEON_NIWO_Litter_mass_trap_Processed.csv", stringsAsFactors = TRUE)
```

2. Make sure R is reading dates as date format; if not change the format to date.

```{r}
#1 
PeterPaul.chem.nutrients$sampledate <- as.Date(PeterPaul.chem.nutrients$sampledate, format = "%Y-%m-%d")

PeterPaul.chem.nutrients.gathered$sampledate <- as.Date(PeterPaul.chem.nutrients.gathered$sampledate, format = "%Y-%m-%d")

Niwot_Ridge.sum$collectDate<-as.Date(Niwot_Ridge.sum$collectDate, format = "%Y-%m-%d")

```
## Define your theme
3. Build a theme and set it as your default theme.

```{r}
mytheme <- theme_classic(base_size = 13) +
  theme(axis.text = element_text(color = "black"), 
        legend.position = "top") 
theme_set(mytheme)

```

## Create graphs
For numbers 4-7, create ggplot graphs and adjust aesthetics to follow best practices for data visualization. Ensure your theme, color palettes, axes, and additional aesthetics are edited accordingly.

4. [NTL-LTER] Plot total phosphorus (`tp_ug`) by phosphate (`po4`), with separate aesthetics for Peter and Paul lakes. Add a line of best fit and color it black. Adjust your axes to hide extreme values.

```{r}
plot4 <- ggplot(PeterPaul.chem.nutrients, aes(x = po4, y = tp_ug, color=lakename)) +
  geom_point(size=0.4) +
  geom_smooth(method = lm, color="black") +
  labs(y="Total Phosphorus", x="Phosphate", title="Plot Total Phosphorus by Phosphate") +
  xlim(0, 40) +
  ylim(0, 180) 
print(plot4) 

```


5. [NTL-LTER] Make three separate boxplots of (a) temperature, (b) TP, and (c) TN, with month as the x axis and lake as a color aesthetic. Then, create a cowplot that combines the three graphs. Make sure that only one legend is present and that graph axes are aligned.

```{r}
# Box and whiskers plot
plot5.1 <-
  ggplot(PeterPaul.chem.nutrients, aes(x = month, y = temperature_C, color=lakename)) +
  geom_boxplot()
print(plot5.1)

plot5.2 <-
  ggplot(PeterPaul.chem.nutrients, aes(x = month, y = tp_ug, color=lakename)) +
  geom_boxplot()
print(plot5.2)

plot5.3 <-
  ggplot(PeterPaul.chem.nutrients, aes(x = month, y = tn_ug, color=lakename)) +
  geom_boxplot() 
print(plot5.3)

# Make two plots legend disappear
plot5.1.2 <-
  ggplot(PeterPaul.chem.nutrients, aes(x = month, y = temperature_C, color=lakename)) +
  geom_boxplot()+
  theme(legend.position="none")
print(plot5.1.2)

plot5.2.2 <-
  ggplot(PeterPaul.chem.nutrients, aes(x = month, y = tp_ug, color=lakename)) +
  geom_boxplot() +
  theme(legend.position="none")
print(plot5.2.2)

library(cowplot)
plot_grid(plot5.1.2, plot5.2.2, plot5.3, nrow = 1, align = 'h', axis = 'lr')

```

Question: What do you observe about the variables of interest over seasons and between lakes? 

> Answer: from the means and distribution of obversations, Peter lake seems to have a relatively lower temperature, higher phosphorus, and higher phosphate than Paul Lake over season. 

6. [Niwot Ridge] Plot a subset of the litter dataset by displaying only the "Needles" functional group. Plot the dry mass of needle litter by date and separate by NLCD class with a color aesthetic. (no need to adjust the name of each land use)
```{r}
plot6<-
  ggplot(subset(Niwot_Ridge.sum, functionalGroup="Needles"), aes(x=collectDate, y=dryMass, color=nlcdClass))+
  geom_point()+
  scale_x_date(date_breaks = "4 months", date_labels = "%b %y")+
  labs(y="Dry Mass", x="Data Collected Date ") +
  mytheme
print(plot6)
```


7. [Niwot Ridge] Now, plot the same plot but with NLCD classes separated into three facets rather than separated by color. 

```{r}
plot7<-
  ggplot(subset(Niwot_Ridge.sum, functionalGroup="Needles"), aes(x=collectDate, y=dryMass))+
  geom_point(size=0.8)+
  facet_wrap(vars(nlcdClass), nrow=3)+
  scale_x_date(date_breaks = "4 months", date_labels = "%b %y")+
  labs(y="Dry Mass", x="Data Collected Date ") 
print(plot7)
```
Question: Which of these plots (6 vs. 7) do you think is more effective, and why? 

> Answer: I think plot 7 is more effective since it compares the amount of dry mass among three land use groups, thus it display more information than plot 6. 
